Visual diagnostics for constrained optimisation with application to guided tours

### 2020-10-16 paper/paper.Rmd
Small changes in the introduction.

- name: Author 1
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of YYY, University of XXX
- name: Author 2
  affiliation: Department of ZZZ, University of WWW
keywords:
- optimisation
- projection pursuit 
- guided tourr
- visual 
- diagnostics 
- R
abstract: Friedman & Tukey commented on their initial paper on projection pursuit in 1974 that "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedures. In this paper, we illustrates the diagnostics of optimisers for projection pursuit guided tour via the construction of a data object and producing diagnostic plots. These diagnostics workflow can be applied generally on optimisers to assess its performance. An R package, ferrn, has been created to implement the diagnostics.
bibliography: biblio.bib
preamble: >
  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs}
  \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 3
editor_options: 
  chunk_output_type: console
---

@@

Visualisation has been widely used in exploratory data analysis. Presenting information in a graphical format often allows people to see information they would otherwise not see. This motivates our work of creating plots to diagnose optimisation algorithms with the aim to understand and compare features of different algorithms.

In an optimization problem the goal is to find the best solution within the space of all feasible solutions, which typically is represented by a set of constraints. The problem consists in optimizing an objective function $f: S \rightarrow \mathbb{R}$ with $S \in \mathbb{R}^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.

The problem context of the optimisation that has been worked in this paper is projection pursuit guided tour, which are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space. Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.

The remainder of the paper is organised as follows. 
Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods. 
Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem and its features, and introduces three main existing algorithms.
Section \ref{vis-diag} presents new visual diagnostics designs. Data object is first defined to pinned down the data structure, several diagnostic plots are then defined with sample examples given.
Section \ref{application} shows the application of how the diagnostic plots designed previously can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.

>>>>

- name: Author 1
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of YYY, University of XXX
- name: Author 2
  affiliation: Department of ZZZ, University of WWW
keywords:
- optimisation
- projection pursuit 
- guided tourr
- visual 
- diagnostics 
- R
abstract: Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function. The initial paper by Friedman & Tukey in 1974 stated that  "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures. In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour. These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance. An R package, ferrn, has been created to implement the diagnostics.
bibliography: biblio.bib
preamble: >
  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs}
  \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 3
editor_options: 
  chunk_output_type: console
---

@@

Visualisation is widely used in exploratory data analysis. Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of. This motivates our work of creating plots to diagnose optimisation algorithms: we aim to understand and compare the behaviour of different algorithms.

The goal of optimisation is to find the best solution within the space of all feasible solutions, typically represented by a set of constraints. The problem can be formulated as the maximisation (or minimisation) of an objective function $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.

The focus of this paper is on optimisation in the projection pursuit guided tour, which is an exploratory data analysis tool that detects interesting structure in high-dimensional data visible in low-dimensional projections. The target of the optimisation is thus to identify the most interesting low-dimensional view, given by a corresponding projection matrix.

The remainder of the paper is organised as follows. 
Section \ref{optim} provides an overview of optimisation methods, specifically line search methods. 
Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and introduces three existing algorithms.
Section \ref{vis-diag} presents the new visual diagnostics. A data structure is defined to capture information during the optimisation, and used in different types of diagnostic plots.
Section \ref{application} shows applications of how these plots can be used to understand and compare different algorithms. We also discuss how these insights contribute to modifications that improve the algorithms.
Finally, Section \ref{implementation} describes the R package: ferrn, that implements the visual diagnostics.
